# R foundations


## Variables & Values

* assign values to variables either with `<-` or `=`
* variable on the left, assigned value on the right
* You decide how to name a variable.
    * Use recognizable names!
    * Don't be lazy. Longer, descriptive names can help immensely.

```{r}
a <- 5
b = 10

a_big_number = 500000
```

### Value Types

* How is data represented in R?
* Types we need right now:
    * numeric
    * character (string / text)
    * logical

```{r, eval = FALSE}
# numeric
1
68542
4.5


# character
"f"
"Hi Mom!"

# logical
TRUE
F

# nothing
NA # there should be something, but it is missing
NULL # truly nothing
```

## Functions

* Everything you do in R is a function!
* Easily identified by function name followed by parenthesis (), e.g. `print()`
* for common functions there are shortcuts


```{r}
# Really everything you do in R is a function!
`<-`(a, 5)
print(a)

# both functions above have shortcuts:
a <- 5
a
```

* most functions need __arguments__
* input, options, parameters to specify what the function should do
* each function has a help page where you can read about arguments


For example, the `cos()` function computes the cosine of a number. The help page specifies what the function expects as an input. You can access the help page with `?cos()` or by searching for the function in the 'Help' tab in Rstudio (bottom right panel).

```{r, error = TRUE}
# the cos function needs the argument x of type numeric
cos(x = 0)

# it throws an error otherwise
cos(x = "3.14")

# argument names can be neglected
cos(1)
```



## Vectors

* most important data structure in R
* building block for more complicated structures
* created with the function `c()`
* think about a container where you can put in multiple things
* vectors are atomic - each entry has to be from the same value type
* each element has an index with which it can be accessed


```{r, fig.cap="R vector model", echo = FALSE}
knitr::include_graphics("assets/r_vector.svg")
```


```{r}
c(23, 1, 60, 21, 21)
```



```{r, error = TRUE}
# vectors are atomic
c(23, 1, 60.5, 21, 21)

# to actually use a vector we have to assign it to a variable
vec = c(23, 1, 60, 21, 21)

vec[1]
vec[3]


# accessing multiple elements at once
vec[1,2]
vec[c(1,2)]
```


```{r, fig.cap="R vector addition", echo = FALSE}
knitr::include_graphics("assets/r_vector_plus.svg")
```

* most of the functions in R are "vectorised"
* i.e. functions are applied to all the elements in a vector


```{r, warning = TRUE}
vec1 = c(23, 1, 60, 21, 21, 5)
vec1 + 6

vec2 = c(5, 1000, 5, 1000, 5, 1000)
vec1 - vec2
```


```{r, fig.cap="R vector multiplication, different length", echo = FALSE}
knitr::include_graphics("assets/r_vector_mul.svg")
```


```{r}
# be careful with vectors of different length!
vec3 = c(1, 2, 1)
vec1 * vec3

vec4 = c(1, 2, 1, 2)
vec1 * vec4
```



## TASK: Vectors and functions

The vector below are the temperature values for 2021-06-02 from the Aasee dataset.
Copy it into an empty are script. 

```{r, results = "hide"}
temperature <- c(18.81,18.73,18.92,18.87,18.76,18.79,18.87,18.84,18.89,18.89,
                 18.93,18.95,18.92,18.97,19.00,18.99,19.00,19.00,18.99,19.00,
                 18.98,18.97,18.97,18.97,18.97,18.98,18.96,18.94,18.94,18.91,
                 18.90,18.89,18.87,18.85,18.85,18.84,18.83,18.81,18.81,18.81,
                 18.81,18.81,18.80,18.79,18.80,18.79,18.79,18.78,18.77,18.78,
                 18.79,18.79,18.79,18.79,18.79,18.79,18.80,18.79,18.81,18.80,
                 18.81,18.81,18.82,18.84,18.87,18.89,18.89,18.91,18.96,19.02,
                 19.11,19.15,18.95,18.92,18.84,18.85,18.96,18.92,18.95,18.95,
                 18.91,19.00,19.02,19.06,19.05,19.47,19.32,19.05,19.66,19.57,
                 19.23,19.42,19.36,19.66,19.79,19.81,19.99,21.10,21.94,22.52,
                 22.81,22.43,22.16,21.55,21.80,21.02,21.85,21.45,22.48,22.73,
                 22.61,22.83,20.85,20.68,19.76,19.81,19.74,19.58,19.35,19.40,
                 19.75,19.76,19.85,19.82,19.86,20.08,20.06,19.85,20.04,19.90,
                 19.96,20.12,19.94,20.20,20.17,20.31,20.17,20.37,20.36,20.13)
```


To get familiar with the R syntax and vectors, do the following tasks:

* Convert the temperature from Celsius to Fahrenheit.
* Try out the following functions: `length()`, `range()`, `seq()`

> Come up with a reasonable questions that someone might have looking at this data.

