# Vectors

* most important data structure in R
* building block for more complicated structures
* created with the function `c()`
* think about a container where you can put in multiple things
* vectors are atomic - each entry has to be from the same value type
* each element has an index with which it can be accessed


```{r, fig.cap="R vector model", echo = FALSE}
knitr::include_graphics("assets/r_vector.svg")
```


```{r}
c(23, 1, 60, 21, 21)
```



```{r, error = TRUE}
# vectors are atomic
c(23, 1, 60.5, 21, 21)

# to actually use a vector we have to assign it to a variable
vec = c(23, 1, 60, 21, 21)

```


## Vector Maths


```{r, fig.cap="R vector addition", echo = FALSE}
knitr::include_graphics("assets/r_vector_plus.svg")
```

* most of the functions in R are "vectorised"
* i.e. functions are applied to all the elements in a vector


```{r, warning = TRUE}
vec1 = c(23, 1, 60, 21, 21, 5)
vec1 + 6

vec2 = c(5, 1000, 5, 1000, 5, 1000)
vec1 - vec2
```


```{r, fig.cap="R vector multiplication, different length", echo = FALSE}
knitr::include_graphics("assets/r_vector_mul.svg")
```


```{r}
# be careful with vectors of different length!
vec3 = c(1, 2, 1)
vec1 * vec3

vec4 = c(1, 2, 1, 2)
vec1 * vec4
```



## Vectors and functions

The vector below are the temperature values for the Aasee in Münster from the 2021-06-02.

```{r, results = "hide"}
temperature <- c(18.81,18.73,18.92,18.87,18.76,18.79,18.87,18.84,18.89,18.89,
                 18.93,18.95,18.92,18.97,19.00,18.99,19.00,19.00,18.99,19.00,
                 18.98,18.97,18.97,18.97,18.97,18.98,18.96,18.94,18.94,18.91,
                 18.90,18.89,18.87,18.85,18.85,18.84,18.83,18.81,18.81,18.81,
                 18.81,18.81,18.80,18.79,18.80,18.79,18.79,18.78,18.77,18.78,
                 18.79,18.79,18.79,18.79,18.79,18.79,18.80,18.79,18.81,18.80,
                 18.81,18.81,18.82,18.84,18.87,18.89,18.89,18.91,18.96,19.02,
                 19.11,19.15,18.95,18.92,18.84,18.85,18.96,18.92,18.95,18.95,
                 18.91,19.00,19.02,19.06,19.05,19.47,19.32,19.05,19.66,19.57,
                 19.23,19.42,19.36,19.66,19.79,19.81,19.99,21.10,21.94,22.52,
                 22.81,22.43,22.16,21.55,21.80,21.02,21.85,21.45,22.48,22.73,
                 22.61,22.83,20.85,20.68,19.76,19.81,19.74,19.58,19.35,19.40,
                 19.75,19.76,19.85,19.82,19.86,20.08,20.06,19.85,20.04,19.90,
                 19.96,20.12,19.94,20.20,20.17,20.31,20.17,20.37,20.36,20.13)
```


## Subsetting

For many questions it is often required to access only parts of the data.
This is one of the most common things you have to deal with in R.


### Indexing

To access single values of a vector, use square brackets `[]`.


```{r, error = TRUE}
# get the third value of the vector
temperature[3]


# accessing multiple elements at once
temperature[1,2]
temperature[c(1,2)]
```

> Using the functions above, what is an easy way to access the first 10 elements of the vector?


### Logical Operators


```{r, fig.cap="R vector logical operation", echo = FALSE}
knitr::include_graphics("assets/r_vector_bigger.svg")
```


A more powerful way to subset a vector is by using logical operators. These functions usually return a boolean object (`TRUE, FALSE`) which can then be used to access certain values.

```{r}
1 > 2
1 < 2

1 == 2

max(temperature) > 22

temperature > 20
hot = temperature > 20

temperature[hot]

```


