# String handling

Modifying strings (i.e. text) is a very common task but can be a bit tricky if you start out.
Here are some functions to try out:

* create strings from multiple "substrings" with the `paste` or `paste0` function

```{r}
n = 5
paste("The number five:", n)


# again functions are vectorized
paste(seq(10), "is", c("odd", "even"))

# paste0 defaults to no character for separating the strings
paste0(n, "five", n)

```


## The stringr package

For advanced string handling we will make our lives easier by using the `stringr` package.

To use functions from a package we have to load it first with the `library` function. The first time you want to use a package you might have to install it first with `install.packages`.


```{r, include = FALSE}
library(stringr)
```




```{r, eval = FALSE}
install.packages("stringr")

library(stringr)
```


Many packages have so called `vignettes` which are tutorial style help pages about the package.
You can access vignettes via the "Packages" Pane in R studio or with a quick [Google search](https://www.google.com/search?q=stringr+vignette).

It is a good idea to go over the vignette when using a new package.


```{r}

s = "Hi Mom!"

# you can indicate in your script from which package a function comes from
# by using ::
stringr::str_length(s)

str_sub(string = s, start = 1, end = 2)


string_vector = paste(seq(10), "is", c("odd", "even"))
string_vector

str_remove(string_vector, pattern = "is ")
str_replace(string_vector, pattern = " is", replacement = ":")

```


## Leading Zeros

`str_pad` one can be quite important to keep you sane. It formats strings to a consistent length by adding the character specified with the `pad` argument.


```{r}
string_vector = paste(seq(15), "is", c("odd", "even"))
string_vector

string_vector_pad = paste(str_pad(seq(15), width = 2, pad = 0, side = "left"),
                          "is",
                          c("odd", "even"))
string_vector_pad
```



        
