# The Aasee Dataset 

To learn the basics about data types, data handling and the fundamentals of R, we will work with a dataset from the Aasee in Muenster:
Inform yourself real quick [here](https://opendata.stadt-muenster.de/dataset/aasee-m%C3%BCnster-tagesaktuelle-sensordaten) about the origin of the dataset. You can download it here: [https://github.com/od-ms/aasee-monitoring](https://github.com/od-ms/aasee-monitoring)

Here is an excerpt of the data:


```{r, echo = FALSE}
df <- read.csv(file = "data/2021-06_aasee.csv")
knitr::kable(
  df[1:10,], caption = 'First entries of the Aasee monitoring dataset',
  booktabs = TRUE
)
```


As a first intermediate goal we want to deal with the following task. We will learn all the necessary skills in R and a bit of statistics in the next lessons, but it is valuable to think about the task / question beforehand! 

> Compare the water temperature of 2021 June 02 and July 21.

* What does this task actually mean?
* Reformulate it as a question
* Outline steps to answer the question

## Solution

<details>
<summary>Click for Answer</summary>


```{r}
# load needed packages
library(stringr)

# data input

june = read.csv("data/2021-06_aasee.csv")
july = read.csv("data/2021-07_aasee.csv")


# actual processing ----
# get the day information from the Datum string
june$day = str_sub(june$Datum, start = 6, end = 10)
july$day = str_sub(july$Datum, start = 6, end = 10)

# reduce data to the needed days
june = june[june$day == "06-02",]
july = july[july$day == "07-21",]

mean(june$Wassertemperatur)
mean(july$Wassertemperatur)

# visual comparison
aasee = rbind(june, july)
boxplot(aasee$Wassertemperatur ~ aasee$day)
```



> What are the flaws of this code?

If we want a different month or day we have to make changes in the code. 
As an R beginner, this is not something you should worry about to much - better spend your time to figure out what the R functions do and get the job done. Later on in your R journey this is something you should keep in mind and write more generalized code by writing your own R functions.


**Side Note: Creating R functions**


```{r}
aasee_daily_mean = function(data, day){
    data$day = str_sub(data$Datum, start = 6, end = 10)
    data = data[data$day == day,]
    result = mean(data$Wassertemperatur)
    return(result)
}



july = read.csv("data/2021-07_aasee.csv")
aasee_daily_mean(july, "07-17")
```


</details>




