# R04: Logical Operators

Logical Operators are essential to efficiently work with R. It is a good idea to spent some time with them and think about their possibilities and usage.

But first: what are logical operators? They are used to compare values. The most intuitive ones are the `<` and `>` operators.

```{r}
4 < 5

6 > 2
```

As you see, both statements return the value `TRUE` because, ...well... 4 is smaller than 5 and 6 is larger then 2. As you might guess, if there is `TRUE` there should also be `FALSE`.


```{r}
24 > 50
```

Logical operators will *always* return `TRUE` or `FALSE`, so the statements you make with them are always questions which you can answer as `TRUE` or `FALSE`. As a data type, this is called a `logical` (or sometimes `boolean value` after the British mathematician George Boole).


```{r}
bool <- 24 < 50
class(bool)
```


To ask the question if two values are the same we use the operator `==`:

```{r}
1.9999999 == 2

a = 5 # one '=' means assignment
a == 5 # two '==' is the logical operator for "is equal"
```



## Logical operators and vectors

Logical operators are vectorized. This means that when we ask the question "Is the vector larger than 10?", we get as many `TRUE` and `FALSE` answers as there are elements in the vector.

```{r, fig.cap="R vector logical operation", echo = FALSE}
knitr::include_graphics("assets/r_vector_bigger.svg")
```


```{r}
vec = c(23, 1, 60, 21, 21, 5)
vec > 10
```

If we safe these answers in the a new variable, we can use the result to subset the vector.
So the questions "Is the vector larger than 10?" serves as an important step to do the operation "Give me all the numbers of this vector that are larger than 10!".

```{r}
large = vec > 10
vec[large]

# in short, we combine both steps to
vec[vec > 10]
```

