# R04: Dataframes


> From a data science perspective, the most important class of objects is the data frame - @Chambers2020


* Dataframes combine multiple vector in a structured format
* Intuitive - most people have seen a spreadsheet before
* Each column is regarded as a vector
* Each vector can have a different data type


```{r, echo = FALSE, fig.cap = "data.frames are a named list of vectors"}
knitr::include_graphics("assets/r_dataframe.svg")
```


* Create one with `data.frame`
* The functions need the column name followed by the vector of values


```{r}
df = data.frame(plotID = seq(3),
                soil_ph = c(5.5, 5.4, 6.1),
                soil_temperature = c(10, 11, 12),
                forest_type = c("coniferous", "coniferous", "deciduous"))

df
```






```{r}
data = read.csv(file = "data/2021-06_aasee.csv", dec = ".", sep = ",")


# show the first few rows of the df
head(data)

# show the last few rows of the df
tail(data)

# get a short summary of the structure
str(data)
```


