# R08: Formula notation

Some functions use a so called `formula` as an argument. Look e.g. at the help page of `boxplot()` to see that it's able to takes an argument `formula` the description

> a formula, such as y ~ grp, where y is a numeric vector of data values to be split into groups according to the grouping variable grp (usually a factor).


```{r}
a = c(1,5,3)
b = c(6,8,7)

a ~ b
class(a~b)
```


A lot of statistical functions use this notation. As a beginner you most likely find formulas in `boxplot()`, `plot()` and `lm()`.

```{r}
data(iris)
boxplot(iris$Petal.Length ~ iris$Species)
```


Think about the `~` like _as a function of_. So `boxplot(iris$Petal.Length ~ iris$Species)` means: Give me boxplots of the petal length as a function of species.


```{r}
# Show me the Sepal.Length as a function of Petal.Length
plot(iris$Sepal.Length ~ iris$Petal.Length)
```



```{r}
# calculate the relation between Sepal.Length and Petal.Length
# model Sepal.Length as a function of Petal.Length


lmod = lm(iris$Sepal.Length ~ iris$Petal.Length)

plot(iris$Sepal.Length ~ iris$Petal.Length)
abline(lmod)
```


