# Formula notation

Some functions in R use a so called `formula` as their argument. Look e.g. at the help page of `boxplot()` to see that it's able to takes an argument `formula` which is describes as follows:

> a formula, such as y ~ grp, where y is a numeric vector of data values to be split into groups according to the grouping variable grp (usually a factor).

So the example code below using the `iris` dataset reads like: make me a boxplot of the petal length split into groups by the species. 

```{r}
data(iris)
boxplot(iris$Petal.Length ~ iris$Species)
```


A lot of statistical functions use this notation. As a beginner you most likely find formulas in `boxplot()`, `plot()` and `lm()`. More general, think about the `~` like _as a function of_. So `boxplot(iris$Petal.Length ~ iris$Species)` means: Give me boxplots of the petal length as a function of the species. Some more examples:


```{r}
# Show me the Sepal.Length as a function of Petal.Length
plot(iris$Sepal.Length ~ iris$Petal.Length)
```



```{r}
# calculate the relation between Sepal.Length and Petal.Length
# model Sepal.Length as a function of Petal.Length


lmod = lm(iris$Sepal.Length ~ iris$Petal.Length)

plot(iris$Sepal.Length ~ iris$Petal.Length)
abline(lmod)
```


